---
title: Face shape sexual dimorphism population comparison
output:
  html_document: default
---

In this notebook we'll compare face shape sexual dimorphism between populations (for magnitude and pattern).
We will use two complementary approaches:
1. First, we will traverse the dendrogram from fineStructure by testing the sex by population interaction term using a MANOVA. 
To do that, we will arbitrarily cut the dendrogram at height 6.5 which results in a total of 11 leaves.
For each of the 11 leaves we used a MANOVA to test for the sex by population interaction term. 
If a non-significant term was found, the leaves were merged.
Finally, the post-hoc tests for pattern and magnitude were used for those branches with significant differences.
We keep track of the initial cluster (11 leaves), and the final cluster (N branches) and export the final dataset.

2. Second, we will compare populations based on their Admixture proportions, with $k = 6$. 
Specifically, in a first attempt we will generate homogeneous populations that have more than 0.7 assignment proportion for any cluster. 
In another attempt we will assign individuals to clusters from which they have most of their ancestry proportions assigned to.
Finally, we will keep track of both attempt and include them into the final dataset.

## Preliminaries

Load libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(car)
library(dendextend)
library(ape)
source("PhenotypicChangeTests.R")
source("FinestructureLibrary.R") # read in the R functions, which also calls the needed packages
```

Read databases

```{r databases, warning=FALSE, message=FALSE}
setwd('..')
path <- getwd()
setwd(paste(path, "/DataBases/Genotypes/Admixture", sep = ""))

#Reading Admixture run
adk6  <- read_delim("geno.6.Q", delim = " ", col_names = FALSE)
adfam <- read_delim("geno.fam", delim = " ", col_names = c("FID", "ID","P","M","Ph")) %>% select(ID)

#Reading covariates
setwd(paste(path, "/DataBases", sep = ""))
covs  <- read_csv("Covariates.csv", col_types = cols(ID = col_character()))

#Reading face PCA
setwd(paste(path, "/DataBases/FacePCA", sep = ""))
scores       <- read_csv("scores.csv", col_names = FALSE)
eigenvectors <- read_csv("eigenvectors.csv", col_names = FALSE)
means        <- read_csv("means.txt", col_names = FALSE)
colnames(scores) <- paste("PC", 1:length(scores), sep="")
face_ids <- read_csv("landmark_ids.txt", col_names = "ID", col_types = cols(ID = col_character()))
face_pca <- cbind(face_ids, scores)

#Reading list of common IDs
setwd(paste(path, "/Results", sep = ""))
common_ids <- read_delim("common_ids.txt", delim = " ", col_names = "ID", col_types = cols(ID = col_character()))

#get only adapt ids
keep  <- which(adfam$ID %in% common_ids$ID)
adk6  <- adk6[keep,]
adfam <- adfam[keep,]

#Subsetting the other databases
tot_samples <- adfam %>% left_join(covs, by = "ID") %>% mutate(BMI = Weight / (Height/100)^2) %>% 
  left_join(face_pca, by = "ID") %>% mutate(Sex = as.factor(Sex))
```

Read and plot the fineStructure dendrogram

```{r tree, warning=FALSE, message=FALSE}
setwd('..')
path <- getwd()
setwd(paste(path, "/DataBases/Genotypes/FineStructure", sep = ""))

treexml <- xmlTreeParse("total_fs_linked.greedy_outputfile.xml") ## read the tree as xml format
ttree   <- extractTree(treexml) ## extract the tree into ape's phylo format
## Will will instead remove "perfect" node labels
ttree$node.label[ttree$node.label=="1"] <- ""
ttree$node.label[ttree$node.label!=""] <- format(as.numeric(ttree$node.label[ttree$node.label!=""]),digits=2)
tdend <- myapetodend(ttree,factor=1) # convert to dendrogram format

#Get number of IDs per pop
mapstate     <- extractValue(treexml,"Pop") # map state as a finestructure clustering
mapstatelist <- popAsList(mapstate) # .. and as a list of individuals in populations
print('Number of pops is')
print(length(mapstatelist))
print('Number of IDs per pop ranges from')
summary(unlist(lapply(mapstatelist, function(x) length(x))))

#Cut
tdend_cut <- cut(tdend, h = 6.5)$upper
tdend_cut_low <- cut(tdend, h = 6.5)$lower
tdend_cut %>% hang.dendrogram %>% plot

fs_groups <- as.data.frame(matrix(NA, nrow = length(labels(tdend)), ncol = 2))
colnames(fs_groups) <- c("ID", "cluster")
fs_groups$ID <- labels(tdend)
ngroups <- max(as.integer(labels(tdend_cut_low)))

for(i in 1:ngroups){
  fs_groups$cluster[fs_groups$ID %in% labels(tdend_cut_low[[i]])] <- i
}

#get only adapt ids
keep       <- which(fs_groups$ID %in% common_ids$ID)
fs_groups  <- fs_groups[keep,]
#keep order of adfam
fs_groups  <- adfam %>% left_join(fs_groups, by = "ID")
summary(as.factor(fs_groups$cluster))

#Adding fs_group initial cluster to tot_samples object
tot_samples <- tot_samples %>% left_join(fs_groups, by = "ID")
colnames(tot_samples)[length(tot_samples)] <- c("cluster_initial")
```

We'll export the truncated dendrogram to newick format for later use

```{r export tree}
tree_cut <- as.phylo(tdend_cut)
setwd(paste(path, "/Results", sep = ""))
write.tree(phy=tree_cut, file="tree_cut.newick")
```

## Analysis 

Traversing the dendrogram. Because 10 independent tests need to be made, we will consider alpha = `r 0.05/10`

```{r tree comparison}
matrix <- tot_samples %>% select(starts_with("PC"))
covs   <- tot_samples %>% select(c("Age", "Height", "BMI","Sex", "cluster_initial"))
covs[,c(1:3)] <- scale(covs[,c(1:3)])

run_Manova <- function(keep){
  fit  <- lm(as.matrix(matrix[keep,]) ~ covs$Age[keep] + covs$Height[keep] + covs$BMI[keep] + 
            covs$Sex[keep] * as.factor(covs$cluster_initial[keep]))
  Manova(fit, type = 3, test.statistic = "Wilks")
}

#Getting model matrix, and runnning vector change
keep <- covs$cluster_initial == 1 | covs$cluster_initial == 2 
run_Manova(keep)
#Changing value of branch 2 to 1
covs$cluster_initial[covs$cluster_initial == 2] <- 1

#Branch1vs3
keep <- covs$cluster_initial == 1 | covs$cluster_initial == 3 
run_Manova(keep)
#Changing value of branch 3 to 1
covs$cluster_initial[covs$cluster_initial == 3] <- 1

#Branch4vs5
keep <- covs$cluster_initial == 4 | covs$cluster_initial == 5 
run_Manova(keep)
#Changing value of branch 5 to 4
covs$cluster_initial[covs$cluster_initial == 5] <- 4

#Branch1vs4
keep <- covs$cluster_initial == 1 | covs$cluster_initial == 4 
run_Manova(keep)
#Changing value of branch 5 to 4
covs$cluster_initial[covs$cluster_initial == 4] <- 1

#Branch6vs7
keep <- covs$cluster_initial == 6 | covs$cluster_initial == 7
run_Manova(keep)
#Changing value of branch 7 to 6
covs$cluster_initial[covs$cluster_initial == 7] <- 6

#Branch6vs8
keep <- covs$cluster_initial == 6 | covs$cluster_initial == 8
run_Manova(keep)
#Changing value of branch 8 to 6
covs$cluster_initial[covs$cluster_initial == 8] <- 6

#Branch9vs10
keep <- covs$cluster_initial == 9 | covs$cluster_initial == 10
run_Manova(keep)
#Changing value of branch 10 to 9
covs$cluster_initial[covs$cluster_initial == 10] <- 9

#Branch 6 vs 9
keep <- covs$cluster_initial == 6 | covs$cluster_initial == 9
run_Manova(keep)
#Changing value of branch 9 to 6
covs$cluster_initial[covs$cluster_initial == 9] <- 6

#Branch 6 vs 11
keep <- covs$cluster_initial == 6 | covs$cluster_initial == 11
run_Manova(keep)

#Branch 1 vs 6
keep <- covs$cluster_initial == 1 | covs$cluster_initial == 6
run_Manova(keep)

#Branch 1 vs 11
keep <- covs$cluster_initial == 1 | covs$cluster_initial == 11
run_Manova(keep)

model_mat  <- model.matrix( ~ Age + Height + BMI + Sex * as.factor(covs$cluster_initial), data=covs)
nonallo_covs_mat <- model_mat[,1:4]
total_covs_mat   <- model_mat[,c(1:2,4)]
ngroups  <- length(levels(as.factor(covs$cluster_initial )))
red_end  <- 5 + (ngroups - 1)
full_end <- 5 + ( (ngroups - 1) * 2)
x_mat_red  <- model_mat[,5:red_end]
x_mat_full <- model_mat[,5:full_end]

nonallo_out_mat <- run_vector_change(matrix, nonallo_covs_mat, x_mat_red, x_mat_full, permute=9999)
total_out_mat   <- run_vector_change(matrix, total_covs_mat, x_mat_red, x_mat_full, permute=9999)

nonallo_angle_table <- pairwise_pvals(nonallo_out_mat)
pvals <- nonallo_angle_table[upper.tri(nonallo_angle_table)]
print("Non allometric angles")
nonallo_angle_table
cat("\n")
nonallo_diff_table  <- pairwise_pvals(nonallo_out_mat, 0)
print("Non allometric magnitues")
nonallo_diff_table
cat("\n")

total_angle_table <- pairwise_pvals(total_out_mat)
pvals <- c(pvals, total_angle_table[upper.tri(total_angle_table)])
print("Total angles")
total_angle_table
cat("\n")
total_diff_table  <- pairwise_pvals(total_out_mat, 0)
print("Total magnitudes")
total_diff_table

```

Number of individuals per final branch

```{r n of branches}
covs <- covs %>% mutate(cluster_final = factor(cluster_initial, labels = c(1,2,3)) )
summary(covs$cluster_final)
cat("\n")
summary(interaction(covs$cluster_final, as.factor(covs$Sex)))

#Adding final cluster to tot_samples object
tot_samples <- tot_samples %>% bind_cols(covs %>% select(cluster_final))
```

## Admixture analysis

### Homogeneous pops

Getting individuals for homogeneous pops

```{r homogenous pops}
h_group <- rep(NA, length(common_ids$ID))
for( i in 1:6){
  h_group[adk6[,i] > 0.7] <- i 
}
summary(as.factor(h_group))
```

We'll use only pops 1, 2, 5, and 6, because of the small sample sizes in the other 2. 

```{r}
#Getting subset of whole sample from groups 1, 2, 5, and 6
tot_samples <- bind_cols(tot_samples, as.tibble(h_group))
colnames(tot_samples)[length(tot_samples)] <- "h_group"
tot_samples$h_group[ tot_samples$h_group == 3 | tot_samples$h_group == 4] <- NA

#Summaries by Sex x Pop
tot_samples %>% select(h_group, Sex, Age, Height, Weight, BMI) %>% group_by(Sex, h_group) %>%
  summarise_all(funs(mean))

#Sample sizes
summary(interaction(tot_samples$h_group, as.factor(tot_samples$Sex)))

```

```{r h pop comparison}
#Getting response and predictor matrices
matrix <- tot_samples %>% filter(!is.na(h_group)) %>% select(starts_with("PC"))
covs   <- tot_samples %>% filter(!is.na(h_group)) %>% select(c("Age", "Height", "BMI","Sex", "h_group"))
covs[,c(1:3)] <- scale(covs[,c(1:3)])

#Getting model matrix, and runnning vector change
model_mat  <- model.matrix( ~ Age + Height + BMI + Sex * as.factor(h_group), data=covs)
nonallo_covs_mat <- model_mat[,1:4]
total_covs_mat   <- model_mat[,c(1:2,4)]
x_mat_red  <- model_mat[,5:8]
x_mat_full <- model_mat[,5:11]
nonallo_out_mat <- run_vector_change(matrix, nonallo_covs_mat, x_mat_red, x_mat_full, permute=9999)
total_out_mat   <- run_vector_change(matrix, total_covs_mat, x_mat_red, x_mat_full, permute=9999)

#Printing pairwise tables
nonallo_angle_table <- pairwise_pvals(nonallo_out_mat)
pvals <- nonallo_angle_table[upper.tri(nonallo_angle_table)]
print("Non allometric angles")
nonallo_angle_table
cat("\n")
nonallo_diff_table  <- pairwise_pvals(nonallo_out_mat, 0)
print("Non allometric magnitudes")
nonallo_diff_table
cat("\n")

total_angle_table <- pairwise_pvals(total_out_mat)
pvals <- c(pvals, total_angle_table[upper.tri(total_angle_table)])
print("Total angles")
total_angle_table
cat("\n")
total_diff_table  <- pairwise_pvals(total_out_mat, 0)
print("Total magnitudes")
total_diff_table
```

### Dividing our entire sample

```{r}
#Create grouping variable t_group, containg the cluster that each individuals has more of their ancestry from
t_group <- apply(adk6, 1, function(x) which.max(x))
summary(as.factor(t_group))

#Getting subset of whole sample from groups 
tot_samples <- bind_cols(tot_samples, as.tibble(t_group))
colnames(tot_samples)[length(tot_samples)] <- "t_group"

#Sample sizes
summary(interaction(tot_samples$t_group, tot_samples$Sex))
```

```{r t pop comparison}
#Getting response and predictor matrices
matrix <- tot_samples %>% select(starts_with("PC"))
covs   <- tot_samples %>% select(c("Age", "Height", "BMI","Sex", "t_group"))
covs[,c(1:3)] <- scale(covs[,c(1:3)])

#Getting model matrix, and runnning vector change
model_mat  <- model.matrix( ~ Age + Height + BMI + Sex * as.factor(t_group), data=covs)
nonallo_covs_mat <- model_mat[,1:4]
total_covs_mat   <- model_mat[,c(1:2,4)]
x_mat_red  <- model_mat[,5:10]
x_mat_full <- model_mat[,5:15]
nonallo_out_mat <- run_vector_change(matrix, nonallo_covs_mat, x_mat_red, x_mat_full, permute=9999)
total_out_mat   <- run_vector_change(matrix, total_covs_mat, x_mat_red, x_mat_full, permute=9999)

#Printing pairwise tables
nonallo_angle_table <- pairwise_pvals(nonallo_out_mat)
pvals <- c(pvals, nonallo_angle_table[upper.tri(nonallo_angle_table)])
print("Non allometric angles")
nonallo_angle_table
cat("\n")
nonallo_diff_table  <- pairwise_pvals(nonallo_out_mat, 0)
print("Non allometric magnitues")
nonallo_diff_table
cat("\n")

total_angle_table <- pairwise_pvals(total_out_mat)
pvals <- c(pvals, total_angle_table[upper.tri(total_angle_table)])
print("Total angles")
total_angle_table
cat("\n")
total_diff_table  <- pairwise_pvals(total_out_mat, 0)
print("Total magnitudes")
total_diff_table
```

## Saving datasets

```{r saving}
setwd(paste(path, "/Results", sep = ""))
write_csv(tot_samples, "tot_samples.csv")
```
