---
title: Main Results
output:
  html_document: default
---

Here you can see the output from the main results of the paper ""

## Preliminaries

Load libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(car)
library(dendextend)
library(ape)
source("PhenotypicChangeTests.R")
source("FinestructureLibrary.R") # read in the R functions, which also calls the needed packages
```

Read databases

```{r databases, warning=FALSE, message=FALSE}
setwd('..')
path <- getwd()
setwd(paste(path, "/DataBases/Genotypes/Admixture", sep = ""))

#Reading Admixture run
adk7  <- read_delim("geno.7.Q", delim = " ", col_names = FALSE)
adfam <- read_delim("geno.fam", delim = " ", col_names = c("FID", "ID","P","M","Ph")) %>% select(ID)

#Reading covariates
setwd(paste(path, "/DataBases", sep = ""))
covs  <- read_csv("Covariates.csv", col_types = cols(ID = col_character()))

#Reading face PCA
setwd(paste(path, "/DataBases/FacePCA", sep = ""))
scores       <- read_csv("scores.csv", col_names = FALSE)
eigenvectors <- read_csv("eigenvectors.csv", col_names = FALSE)
means        <- read_csv("means.txt", col_names = FALSE)
colnames(scores) <- paste("PC", 1:length(scores), sep="")
face_ids <- read_csv("landmark_ids.txt", col_names = "ID", col_types = cols(ID = col_character()))
face_pca <- cbind(face_ids, scores)

#get only adapt ids
keep  <- which(adfam$ID %in% face_ids$ID)
adk7  <- adk7[keep,]
adfam <- adfam[keep,]

#Subsetting the other databases
tot_samples <- adfam %>% left_join(covs, by = "ID") %>% mutate(BMI = Weight / (Height/100)^2) %>% 
  left_join(face_pca, by = "ID") %>% mutate(Sex = as.factor(Sex)) %>% distinct(ID, .keep_all = TRUE)
```

Read and plot the fineStructure dendrogram

```{r tree, warning=FALSE, message=FALSE}
setwd('..')
path <- getwd()
setwd(paste(path, "/DataBases/Genotypes/FineStructure", sep = ""))

treexml <- xmlTreeParse("total_fs_linked.tree.xml") ## read the tree as xml format
ttree   <- extractTree(treexml) ## extract the tree into ape's phylo format
## Will will instead remove "perfect" node labels
ttree$node.label[ttree$node.label=="1"] <- ""
ttree$node.label[ttree$node.label!=""] <- format(as.numeric(ttree$node.label[ttree$node.label!=""]),digits=2)
tdend <- myapetodend(ttree,factor=1) # convert to dendrogram format

#Get number of IDs per pop
mapstate     <- extractValue(treexml,"Pop") # map state as a finestructure clustering
mapstatelist <- popAsList(mapstate) # .. and as a list of individuals in populations
print('Number of pops is')
print(length(mapstatelist))
print('Number of IDs per pop ranges from')
summary(unlist(lapply(mapstatelist, function(x) length(x))))

#Cut
height_cut <- 900000
tdend_cut <- cut(tdend, h = height_cut)$upper
tdend_cut_low <- cut(tdend, h = height_cut)$lower
tdend_cut %>% hang.dendrogram(hang = -1) %>% plot

fs_groups <- as.data.frame(matrix(NA, nrow = length(labels(tdend)), ncol = 2))
colnames(fs_groups) <- c("ID", "cluster")
fs_groups$ID <- labels(tdend)
ngroups <- max(as.integer(labels(tdend_cut_low)))

for(i in 1:ngroups){
  fs_groups$cluster[fs_groups$ID %in% labels(tdend_cut_low[[i]])] <- i
}

#get only adapt ids
keep       <- which(fs_groups$ID %in% face_ids$ID)
fs_groups  <- fs_groups[keep,]
#keep order of adfam
fs_groups  <- adfam %>% left_join(fs_groups, by = "ID")
summary(as.factor(fs_groups$cluster))

#Adding fs_group initial cluster to tot_samples object
tot_samples <- tot_samples %>% left_join(fs_groups, by = "ID")
colnames(tot_samples)[length(tot_samples)] <- c("cluster_initial")
```

We'll export the truncated dendrogram to newick format for later use

```{r export tree}
tree_cut <- as.phylo(tdend_cut)
plot(tree_cut)
setwd(paste(path, "/Results", sep = ""))
write.tree(phy=tree_cut, file="tree_cut.newick")
```

## Analysis 

### Summary values

From the merge of the three datasets, we ended up with `r dim(tot_samples)[1]` individuals (females = `r summary(as.factor(tot_samples$Sex))[1]`, males = `r summary(as.factor(tot_samples$Sex))[2]` )

```{r summary total}
#summary
tot_samples %>% select(Age, Height, Weight, BMI) %>% summarise_all(funs(mean, sd, min, max))
#summary by sex
tot_samples %>% select(Sex, Age, Height, Weight, BMI) %>% group_by(Sex) %>% 
  summarise_all(funs(mean, sd, min, max))

#Differences between sexes
wilcox.test(tot_samples$Age ~ tot_samples$Sex, conf.int = TRUE)
wilcox.test(tot_samples$Height ~ tot_samples$Sex, conf.int = TRUE)
wilcox.test(tot_samples$Weight ~ tot_samples$Sex, conf.int = TRUE)
wilcox.test(tot_samples$BMI ~ tot_samples$Sex, conf.int = TRUE)
```

Running pairwise comparison through tree. Because 10 independent tests need to be made, we will consider alpha = `r 0.05/10`

```{r tree comparison}
matrix <- tot_samples %>% select(starts_with("PC"))
covs   <- tot_samples %>% select(c("Age", "Height", "BMI","Sex", "cluster_initial"))
covs[,c(1:3)] <- scale(covs[,c(1:3)])

#Branch2vs3
keep <- covs$cluster_initial == 2 | covs$cluster_initial == 3 
fit  <- lm(as.matrix(matrix[keep,]) ~ covs$Age[keep] + covs$Height[keep] + covs$BMI[keep] + 
            covs$Sex[keep] * as.factor(covs$cluster_initial[keep]))
Manova(fit, type = 3, test.statistic = "Wilks")
#Changing value of branch 3 to 2
covs$cluster_initial[covs$cluster_initial == 3] <- 2

#Branch1vs2
keep <- covs$cluster_initial == 1 | covs$cluster_initial == 2 
fit  <- lm(as.matrix(matrix[keep,]) ~ covs$Age[keep] + covs$Height[keep] + covs$BMI[keep] + 
            covs$Sex[keep] * as.factor(covs$cluster_initial[keep]))
Manova(fit, type = 3, test.statistic = "Wilks")
#Changing value of branch 2 to 1
covs$cluster_initial[covs$cluster_initial == 2] <- 1

#Branch5vs6
keep <- covs$cluster_initial == 5 | covs$cluster_initial == 6 
fit  <- lm(as.matrix(matrix[keep,]) ~ covs$Age[keep] + covs$Height[keep] + covs$BMI[keep] + 
            covs$Sex[keep] * as.factor(covs$cluster_initial[keep]))
Manova(fit, type = 3, test.statistic = "Wilks")
#Changing value of branch 6 to 5
covs$cluster_initial[covs$cluster_initial == 6] <- 5

#Branch5vs7
keep <- covs$cluster_initial == 5 | covs$cluster_initial == 7
fit  <- lm(as.matrix(matrix[keep,]) ~ covs$Age[keep] + covs$Height[keep] + covs$BMI[keep] + 
            covs$Sex[keep] * as.factor(covs$cluster_initial[keep]))
Manova(fit, type = 3, test.statistic = "Wilks")
#Changing value of branch 7 to 5
covs$cluster_initial[covs$cluster_initial == 7] <- 5

#Branch4vs5
keep <- covs$cluster_initial == 4 | covs$cluster_initial == 5
fit  <- lm(as.matrix(matrix[keep,]) ~ covs$Age[keep] + covs$Height[keep] + covs$BMI[keep] + 
            covs$Sex[keep] * as.factor(covs$cluster_initial[keep]))
Manova(fit, type = 3, test.statistic = "Wilks")
#Changing value of branch 5 to 4
covs$cluster_initial[covs$cluster_initial == 5] <- 4

#Branch8vs9
keep <- covs$cluster_initial == 8 | covs$cluster_initial == 9
fit  <- lm(as.matrix(matrix[keep,]) ~ covs$Age[keep] + covs$Height[keep] + covs$BMI[keep] + 
            covs$Sex[keep] * as.factor(covs$cluster_initial[keep]))
Manova(fit, type = 3, test.statistic = "Wilks")
#Changing value of branch 9 to 8
covs$cluster_initial[covs$cluster_initial == 9] <- 8

#Branch4vs8
keep <- covs$cluster_initial == 4 | covs$cluster_initial == 8
fit  <- lm(as.matrix(matrix[keep,]) ~ covs$Age[keep] + covs$Height[keep] + covs$BMI[keep] + 
            covs$Sex[keep] * as.factor(covs$cluster_initial[keep]))
Manova(fit, type = 3, test.statistic = "Wilks")
#Changing value of branch 8 to 4
covs$cluster_initial[covs$cluster_initial == 8] <- 4

#Branch4vs10
keep <- covs$cluster_initial == 4 | covs$cluster_initial == 10
fit  <- lm(as.matrix(matrix[keep,]) ~ covs$Age[keep] + covs$Height[keep] + covs$BMI[keep] + 
            covs$Sex[keep] * as.factor(covs$cluster_initial[keep]))
Manova(fit, type = 3, test.statistic = "Wilks")

#Branch1vs4
keep <- covs$cluster_initial == 1 | covs$cluster_initial == 4
fit  <- lm(as.matrix(matrix[keep,]) ~ covs$Age[keep] + covs$Height[keep] + covs$BMI[keep] + 
            covs$Sex[keep] * as.factor(covs$cluster_initial[keep]))
Manova(fit, type = 3, test.statistic = "Wilks")

#Branch1vs10
keep <- covs$cluster_initial == 1 | covs$cluster_initial == 10
fit  <- lm(as.matrix(matrix[keep,]) ~ covs$Age[keep] + covs$Height[keep] + covs$BMI[keep] + 
            covs$Sex[keep] * as.factor(covs$cluster_initial[keep]))
Manova(fit, type = 3, test.statistic = "Wilks")

model_mat  <- model.matrix( ~ Age + Height + BMI + Sex * as.factor(covs$cluster_initial), data=covs)
nonallo_covs_mat <- model_mat[,1:4]
total_covs_mat   <- model_mat[,c(1:2,4)]
ngroups  <- length(levels(as.factor(covs$cluster_initial )))
red_end  <- 5 + (ngroups - 1)
full_end <- 5 + ( (ngroups - 1) * 2)
x_mat_red  <- model_mat[,5:red_end]
x_mat_full <- model_mat[,5:full_end]

nonallo_out_mat <- run_vector_change(matrix, nonallo_covs_mat, x_mat_red, x_mat_full, permute=9999)
total_out_mat   <- run_vector_change(matrix, total_covs_mat, x_mat_red, x_mat_full, permute=9999)

nonallo_angle_table <- pairwise_pvals(nonallo_out_mat)
pvals <- nonallo_angle_table[upper.tri(nonallo_angle_table)]
print("Non allometric angles")
nonallo_angle_table
cat("\n")
nonallo_diff_table  <- pairwise_pvals(nonallo_out_mat, 0)
print("Non allometric magnitues")
nonallo_diff_table
cat("\n")

total_angle_table <- pairwise_pvals(total_out_mat)
pvals <- c(pvals, total_angle_table[upper.tri(total_angle_table)])
print("Total angles")
total_angle_table
cat("\n")
total_diff_table  <- pairwise_pvals(total_out_mat, 0)
print("Total magnitudes")
total_diff_table

```

Number of individuals per final branch

```{r n of branches}
covs <- covs %>% mutate(cluster_final = factor(cluster_initial, labels = c(1,2,3)) )
summary(covs$cluster_final)
cat("\n")
summary(interaction(covs$cluster_final, as.factor(covs$Sex)))

#Adding final cluster to tot_samples object
tot_samples <- tot_samples %>% bind_cols(covs %>% select(cluster_final))
```

## Saving datasets

```{r saving}
setwd(paste(path, "/Results", sep = ""))
write_csv(tot_samples, "tot_samples.csv")
```

## Second Analysis

Here we'll use Admixture values

### Homogeneous pops

Getting individuals for homogeneous pops

```{r homogenous pops}
h_group <- rep(NA, length(tot_samples$ID))
for( i in 1:6){
  h_group[adk7[,i] > 0.7] <- i 
}
summary(as.factor(h_group))
```

We'll use only pops 2, 3, 4, and 5, because of the small sample sizes in the other. 

```{r}
#Getting subset of whole sample from groups 1, 2, 4, and 6
h_comparison <- bind_cols(as_tibble(h_group), tot_samples)
h_comparison$value[ h_comparison$value == 6 ] <- NA
h_comparison <- na.omit(h_comparison)

#Sample sizes
summary(interaction(h_comparison$value, as.factor(h_comparison$Sex)))

table(h_comparison$value, h_comparison$cluster_initial)
table(h_comparison$value, h_comparison$cluster_final)
```

```{r h pop comparison}
#Getting response and predictor matrices
matrix <- h_comparison %>% select(starts_with("PC"))
covs   <- h_comparison %>% select(c("Age", "Height", "BMI","Sex", "value"))
covs[,c(1:3)] <- scale(covs[,c(1:3)])

#Getting model matrix, and runnning vector change
model_mat  <- model.matrix( ~ Age + Height + BMI + Sex * as.factor(value), data=covs)
nonallo_covs_mat <- model_mat[,1:4]
total_covs_mat   <- model_mat[,c(1:2,4)]
x_mat_red  <- model_mat[,5:8]
x_mat_full <- model_mat[,5:11]
nonallo_out_mat <- run_vector_change(matrix, nonallo_covs_mat, x_mat_red, x_mat_full, permute=9999)
total_out_mat   <- run_vector_change(matrix, total_covs_mat, x_mat_red, x_mat_full, permute=9999)

#Printing pairwise tables
nonallo_angle_table <- pairwise_pvals(nonallo_out_mat)
pvals <- nonallo_angle_table[upper.tri(nonallo_angle_table)]
print("Non allometric angles")
nonallo_angle_table
cat("\n")
nonallo_diff_table  <- pairwise_pvals(nonallo_out_mat, 0)
print("Non allometric magnitues")
nonallo_diff_table
cat("\n")

total_angle_table <- pairwise_pvals(total_out_mat)
pvals <- c(pvals, total_angle_table[upper.tri(total_angle_table)])
print("Total angles")
total_angle_table
cat("\n")
total_diff_table  <- pairwise_pvals(total_out_mat, 0)
print("Total magnitudes")
total_diff_table
```

### Dividing our entire sample

```{r}
#Create grouping variable t_group, containg the cluster that each individuals has more of their ancestry from
t_group <- apply(adk7, 1, function(x) which.max(x))
summary(as.factor(t_group))
```

```{r t pop comparison}
#Getting subset of whole sample from groups 
t_comparison <- bind_cols(as.tibble(t_group), tot_samples)
t_comparison <- na.omit(t_comparison)
t_comparison <- t_comparison %>% filter(Sex == "Female" | Sex == "Male") %>% mutate(value = as.factor(value))

#Sample sizes
summary(interaction(t_comparison$value, as.factor(t_comparison$Sex)))

#Getting response and predictor matrices
matrix <- t_comparison %>% select(starts_with("PC"))
covs   <- t_comparison %>% select(c("Age", "Height", "BMI","Sex", "value"))
covs[,c(1:3)] <- scale(covs[,c(1:3)])

#Getting model matrix, and runnning vector change
model_mat  <- model.matrix( ~ Age + Height + BMI + Sex * value, data=covs)
nonallo_covs_mat <- model_mat[,1:4]
total_covs_mat   <- model_mat[,c(1:2,4)]
x_mat_red  <- model_mat[,5:9]
x_mat_full <- model_mat[,5:13]
nonallo_out_mat <- run_vector_change(matrix, nonallo_covs_mat, x_mat_red, x_mat_full, permute=9999)
total_out_mat   <- run_vector_change(matrix, total_covs_mat, x_mat_red, x_mat_full, permute=9999)

#Printing pairwise tables
nonallo_angle_table <- pairwise_pvals(nonallo_out_mat)
pvals <- c(pvals, nonallo_angle_table[upper.tri(nonallo_angle_table)])
print("Non allometric angles")
nonallo_angle_table
cat("\n")
nonallo_diff_table  <- pairwise_pvals(nonallo_out_mat, 0)
print("Non allometric magnitues")
nonallo_diff_table
cat("\n")

total_angle_table <- pairwise_pvals(total_out_mat)
pvals <- c(pvals, total_angle_table[upper.tri(total_angle_table)])
print("Total angles")
total_angle_table
cat("\n")
total_diff_table  <- pairwise_pvals(total_out_mat, 0)
print("Total magnitudes")
total_diff_table
```

## Allometric differences

```{r allo}
library(RRPP)

matrix <- tot_samples %>% select(starts_with("PC"))
covs   <- tot_samples %>% select(c("Age", "Height", "BMI","Sex", "cluster_final"))
covs[,c(1:3)] <- scale(covs[,c(1:3)])

dat <- list(Y = matrix, X = covs)

fit <- lm.rrpp(as.matrix(dat$Y) ~ dat$X$Age + dat$X$BMI + dat$X$Height + 
                 (dat$X$Sex * dat$X$cluster_final) + dat$X$Height * (dat$X$Sex * dat$X$cluster_final), 
               SS.type = "III", data=dat, iter = 9999, print.progress = F)

anova(fit, effect.type = "F")

```
