---
title: Distance Analysis
output:
  html_document: default
---

The following script contains the distance analysis. Specifically, we wanted to know whether differences in face shape sexual dimorphism increase with increasing genetic distance. To do that we ran a Mantel test between the genetic distance (haplotype-based TVD) and the angle between vectors of sex dimorphism.

## Preliminaries

Load libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
#For mantel test
library(ade4)
#For latex
library(Hmisc)
library(RRPP)
source("PhenotypicChangeTests.R")
```

Read databases

```{r databases, warning=FALSE, message=FALSE}
setwd('..')
path <- getwd()

#Reading tot_samples database
setwd(paste(path, "/Results", sep = ""))
tot_samples <- read_csv("tot_samples.csv")
tot_samples <- tot_samples %>% mutate(Sex = as.factor(Sex))

#Reading pop coancestry matrix
setwd(paste(path, "/Results", sep = ""))
tvds         <- read_csv("tvds.csv", col_names = FALSE)
```

## Analysis 

Mantel test

```{r}
#Generating distances from angles between initial leaves
matrix <- tot_samples %>% select(starts_with("PC"))
covs   <- tot_samples %>% select(c("Age", "Height", "BMI","Sex", "cluster_initial"))
covs[,c(1:3)] <- scale(covs[,c(1:3)])

#Getting model matrix, and runnning vector change
model_mat  <- model.matrix( ~ Age + Height + BMI + Sex * as.factor(cluster_initial), data=covs)
nonallo_covs_mat <- model_mat[,1:4]
x_mat_red  <- model_mat[,5:14]
x_mat_full <- model_mat[,5:23]
nonallo_out_mat <- run_vector_change(matrix, nonallo_covs_mat, x_mat_red, x_mat_full, permute=9999)

#Printing pairwise tables
nonallo_angle_table <- as.dist(pairwise_pvals(nonallo_out_mat))

mantel.rtest(as.dist(tvds), nonallo_angle_table, nrepet = 9999)
#plot(mantel.rtest(as.dist(tvds), nonallo_angle_table, nrepet = 9999))
```

Morphological distance

```{r morpho}
y.mat <- as.matrix(select(tot_samples, starts_with("PC") ))
covs  <- select(tot_samples, c("Age", "Height", "BMI", "cluster_initial"))
covs[,c(1:3)] <- scale(covs[,c(1:3)])
covs$cluster_initial <- as.factor(covs$cluster_initial)
model_mat  <- model.matrix( ~ Age + Height + BMI + cluster_initial, data=covs)

b.mat <- solve((t(model_mat) %*% model_mat)) %*% (t(model_mat) %*% y.mat)

base <- c(1, rep(0,3)) # Base contains the intercept and zeros equal to the number of covariates 
base <- t(replicate(10, base))
pops <- rbind(c(0,0,0,0,0,0,0,0,0), 
              c(1,0,0,0,0,0,0,0,0), 
              c(0,1,0,0,0,0,0,0,0), 
              c(0,0,1,0,0,0,0,0,0), 
              c(0,0,0,1,0,0,0,0,0), 
              c(0,0,0,0,1,0,0,0,0),
              c(0,0,0,0,0,1,0,0,0),
              c(0,0,0,0,0,0,1,0,0),
              c(0,0,0,0,0,0,0,1,0),
              c(0,0,0,0,0,0,0,0,1))

x.ls <- cbind(base, pops)
obs.ls <- x.ls %*% b.mat

madist <- matrix(0, nrow = 10, ncol = 10)
testdist <- matrix(0, nrow = 10, ncol = 10)
for(i in 1:9){
  for(l in 2:10){
    if(i<l){
      keepi  <- covs$cluster_initial == i
      keepl  <- covs$cluster_initial == l
      covi   <- cov(y.mat[keepi,])
      covl   <- cov(y.mat[keepl,])
      covpool <- (sum(keepi)*covi + sum(keepl)*covl) / (sum(keepi) + sum(keepl) - 2)
      d <- t(obs.ls[i,] - obs.ls[l,]) %*% solve(covpool) %*% (obs.ls[i,] - obs.ls[l,])
      madist[l,i] <- d 
} } }

mantel.rtest(as.dist(madist), nonallo_angle_table, nrepet = 9999)

```

Latex table

```{r latextab}
tab  <- as.matrix(nonallo_angle_table)
tab2 <- as.matrix(nonallo_angle_table)
tab[lower.tri(tab)]   <- as.matrix(tvds)[lower.tri(as.matrix(tvds))]
tab2[lower.tri(tab2)] <- as.matrix(madist)[lower.tri(as.matrix(madist))]
names <- paste("L", seq(1,10), sep = "")
rownames(tab) <- names
colnames(tab) <- names
diag(tab) <- NA
rownames(tab2) <- names
colnames(tab2) <- names
diag(tab2) <- NA

tab_latex <- latex(round(tab, 3), file="", title="", 
                   caption="Between populations distance matrix. Upper diagonal show the $\\theta$ values, while lower diagonal shows TVDS values", 
                   label = "", booktabs = TRUE)

tab_latex <- latex(round(tab2, 3), file="", title="", 
                   caption="Between populations distance matrix. Upper diagonal show the $\\theta$ values, while lower diagonal shows Mahalanovis $D^{2}$", 
                   label = "", booktabs = TRUE)
```

Export table

```{r exporttable}
setwd(paste(path, "/Results", sep = ""))
write_csv(as.data.frame(tab), "distance_matrices.csv", col_names = FALSE)
write_csv(as.data.frame(tab2), "distance_matrices2.csv", col_names = FALSE)
```
