---
title: Generating Faces
output:
  html_document: default
---

In this script we will generate the faces for a series of figures. This script is non-reproducible due to privacy reasons.

## Preliminaries

Load libraries

```{r libraries}
library(tidyverse)
```

Load databases

```{r databases, message=FALSE}
setwd('..')
path <- getwd()
setwd(paste(path, "/DataBases/FacePCA", sep = ""))

means  <- read_csv("means.txt", col_names = FALSE)
scores <- read_csv("scores.csv", col_names = FALSE)
eigenvec <- read_csv("eigenvectors.csv", col_names = FALSE)

#Reading tot_samples database
setwd(paste(path, "/Results", sep = ""))
tot_samples <- read_csv("tot_samples.csv")
tot_samples <- tot_samples %>% mutate(Sex = as.factor(Sex))
```

## Faces for PCs

Here we will generate the face shape of the projected faces on different axes of the face shape PCA.

```{r pc landmarks}
npcs   <- ncol(scores)
nfaces <- 4 
PCs <- matrix(0, nrow = nfaces * 2, ncol = npcs)
sds <- apply(scores, 2, function(x) sd(x))

l <- 1
for(i in 1:4){
  PCs[l,i] <- sds[i] * 3
  l <- l +1
  PCs[l,i] <- -(sds[i] * 3)
  l <- l +1
}

new_means <- matrix(0, nrow = nfaces * 2, ncol = dim(eigenvec)[1])

for(i in 1:nrow(new_means)){
  new_means[i,] <- t(as.matrix(means))
}

PC_landmarks <- (PCs %*% t(as.matrix(eigenvec)) ) + new_means

#Saving dataset
setwd(paste(path, "/Results/Landmarks", sep = ""))
write_csv(as.data.frame(PC_landmarks), "PC_landmarks.csv", col_names = FALSE)

```

## Sex dimorphism

We'll generate the faces of male and females in the total sample

```{r}
#Exaggerated faces
ex = 1

#Separating datasets
Y <- tot_samples %>% select(starts_with("PC"))
X <- tot_samples %>% select(Age, BMI, Height, Sex)
X[,1:3] <- scale(X[,1:3])

#Getting coefficients
fit <- lm(as.matrix(Y) ~ Age + BMI + Height + Sex, data = X)
avg_female <- fit$coefficients[1,] 
avg_male   <- fit$coefficients[1,] + fit$coefficients[5,]

nrows = 13*2
avg_faces <- data.frame(matrix(0, nrow = nrows, ncol = 13))
avg_faces[25,] <- avg_female
avg_faces[26,] <- avg_male

#Reducing the number of PCs, from 1 to 12
row = 1
for(i in 1:12){
  avg_faces[row,1:i] <- avg_female[1:i]
  row = row + 1
  avg_faces[row,1:i] <- avg_male[1:i]
  row = row + 1
}

avg_faces <- avg_faces * ex

new_means <- matrix(0, nrow = nrows, ncol = dim(eigenvec)[1])
for(i in 1:nrow(new_means)){
  new_means[i,] <- t(as.matrix(means))
}

avg_landmarks <- (as.matrix(avg_faces) %*% t(as.matrix(eigenvec)) ) + new_means

#Saving dataset
setwd(paste(path, "/Results/Landmarks", sep = ""))
write_csv(as.data.frame(avg_landmarks), "avg_landmarks.csv", col_names = FALSE)

```

Here we'll generate the faces of male and female for the 3 final branches, including the common patterns and their difference

```{r sex dimorphism}
#Exaggerated faces
ex = 1

#Separating datasets
Y <- tot_samples %>% select(starts_with("PC"))
X <- tot_samples %>% select(Age, BMI, Height, Sex, cluster_final)
X[,1:3] <- scale(X[,1:3])

#Getting coefficients
fit <- lm(as.matrix(Y) ~ Age + BMI + Height + Sex * as.factor(cluster_final), data = X)

x_ls <- matrix(c(1, 0, 0, 0, 0, 0, 0, 0, 0, 
                 1, 0, 0, 0, 1, 0, 0, 0, 0,
                 1, 0, 0, 0, 0, 1, 0, 0, 0,
                 1, 0, 0, 0, 1, 1, 0, 1, 0,
                 1, 0, 0, 0, 0, 0, 1, 0, 0,
                 1, 0, 0, 0, 1, 0, 1, 0, 1), nrow=6, byrow = T)

obs_ls <- x_ls %*% as.matrix(fit$coefficients)
obs_ls <- obs_ls * ex

obs_ls <- rbind(obs_ls ,(obs_ls[1,] + obs_ls[3,] + obs_ls[5,]) /3)
obs_ls <- rbind(obs_ls, (obs_ls[2,] + obs_ls[4,] + obs_ls[6,]) /3)

for(i in seq(1,6,2)){
  obs_ls <- rbind(obs_ls, ( obs_ls[i,]  - obs_ls[7,]) )
  obs_ls <- rbind(obs_ls, ( obs_ls[i+1,]  - obs_ls[8,]) )
}

new_means <- matrix(0, nrow = nrow(obs_ls), ncol = dim(eigenvec)[1])
for(i in 1:nrow(new_means)){
  new_means[i,] <- t(as.matrix(means))
}

sex_landmarks <- (as.matrix(obs_ls) %*% t(as.matrix(eigenvec)) ) + new_means

#Saving dataset
setwd(paste(path, "/Results/Landmarks", sep = ""))
write_csv(as.data.frame(sex_landmarks), "sex_landmarks.csv", col_names = FALSE)

```


We'll generate the faces of males and females for the 3 final branches, and the 10 initial leaves.

```{r sex dimorphism}
#Define functions
get_landmarks <- function(dat, type = "non-allo", ex = 1){
  #
  #
  
  if(type == "non-allo"){
    fit <- lm(as.matrix(select(dat, starts_with("PC"))) ~ dat$Age + dat$BMI + dat$Sex + dat$Height)
    coeff1  <- fit$coefficients[1,] *  ex
    coeff2  <- ( fit$coefficients[1,] + fit$coefficients[4,] ) * ex
  } else if(type == "total"){
    fit <- lm(as.matrix(select(dat, starts_with("PC"))) ~ dat$Age + dat$BMI + dat$Sex )
    coeff1 <- fit$coefficients[1,] * ex
    coeff2 <- ( fit$coefficients[1,] + fit$coefficients[4,] ) * ex
  } else if(type == "allo"){
    fit <- lm(as.matrix(select(dat, starts_with("PC"))) ~ dat$Age + dat$BMI + dat$Sex + dat$Height)
    #Get difference by sex
    mean_fem  <- dat %>% filter(Sex == "Female") %>% select(Sex, Height) %>% summarise(mean(Height))
    mean_male <- dat %>% filter(Sex == "Male") %>% select(Sex, Height) %>% summarise(mean(Height))
    diff       <- as.numeric(mean_male - mean_fem)
    coeff1 <- ( fit$coefficients[1,] + (fit$coefficients[5,] * as.numeric(mean_fem)) ) * ex
    coeff2 <- ( fit$coefficients[1,] + (fit$coefficients[5,] * as.numeric(mean_male)) ) * ex
  }
  
  coeffs <- rbind(coeff1, coeff2)

  new_means <- matrix(0, nrow = nrow(coeffs), ncol = dim(eigenvec)[1])
  
  for(i in 1:nrow(new_means)){
    new_means[i,] <- t(as.matrix(means))
  }
  
  landmarks <- (as.matrix(coeffs) %*% t(as.matrix(eigenvec)) ) + new_means
  
  return(landmarks)
}

ngroups <- length(levels(as.factor(tot_samples$cluster_final)))
landmark_set <- matrix(0, nrow = ngroups * 3 * 2, ncol = dim(eigenvec)[1])
row = 1
for( i in 1:ngroups){
  dat <- tot_samples %>% filter(cluster_final == i)
  dat[,3:6] <- scale(dat[,3:6])
  lands_tot  <- get_landmarks(dat, "total", ex = 1.5)
  lands_non  <- get_landmarks(dat, "non-allo", ex = 1.5)
  lands_allo <- get_landmarks(dat, "allo", ex = 1.5)
  landmark_set[row,] <- lands_tot[1,]
  row = row + 1
  landmark_set[row,] <- lands_tot[2,]
  row = row + 1
  landmark_set[row,] <- lands_non[1,]
  row = row + 1
  landmark_set[row,] <- lands_non[2,]
  row = row + 1
  landmark_set[row,] <- lands_allo[1,]
  row = row + 1
  landmark_set[row,] <- lands_allo[2,]
  row = row + 1

}


#Saving dataset
setwd(paste(path, "/Results/Landmarks", sep = ""))
write_csv(as.data.frame(landmark_set), "sex_decomp_landmarks.csv", col_names = FALSE)
#write_csv(as.data.frame(sex_3b_landmarks), "sex_3b_landmarks.csv", col_names = FALSE)
#write_csv(as.data.frame(sex_11l_landmarks), "sex_11l_landmarks.csv", col_names = FALSE)
```


Here we will apply the sex dimorphism vectors on the same average face

```{r}

y.mat <- as.matrix(select(tot_samples, starts_with("PC") ))
covs  <- select(tot_samples, c("Age", "Height", "BMI","Sex", "cluster_final"))
covs[,c(1:3)] <- scale(covs[,c(1:3)])
covs$cluster_final <- as.factor(covs$cluster_final)
model_mat  <- model.matrix( ~ Age + Height + BMI + Sex * cluster_final, data=covs)

b.mat <- solve((t(model_mat) %*% model_mat)) %*% (t(model_mat) %*% y.mat)

base <- c(1, rep(0,3)) # Base contains the intercept and zeros equal to the number of covariates 
base <- t(replicate(6, base))
sex_groups <- rbind(c(0,0,0,0,0), c(1,0,0,0,0), c(0,1,0,0,0), c(1,1,0,1,0), c(0,0,1,0,0), c(1,0,1,0,1))

x.ls <- cbind(base, sex_groups)

obs.ls <- x.ls %*% b.mat

#Getting vectors
obs.vect <- matrix(0, nrow = 3 , ncol = ncol(obs.ls)) 
row <- 1
for (i in seq(1,6,2)){
  obs.vect[row,] <- obs.ls[i,] - obs.ls[i+1,] # These are the phenotypic change vectors
  row <- row + 1
}

#get length to set up to unit distance
obs.d <- matrix(0, nrow = 3, ncol = 1)
for (i in 1:3){
  obs.d[i,] <- sqrt(t(obs.vect[i,]) %*% obs.vect[i,]) # These are lengths of vectors
}

obs.vect.unit <- matrix(0, nrow = 3 , ncol = ncol(obs.ls)) 
for (i in 1:3){
  obs.vect.unit[i,] <- obs.vect[i,] / obs.d[i,]
}

average_face <- apply(obs.ls, 2, function(x) mean(x))

sex_vector_average <- matrix(0, nrow = 6, ncol = ncol(obs.ls))
row <- 1
for (i in 1:3){
  sex_vector_average[row, ] <- (average_face + obs.vect[i,] ) * 1.5
  row <- row + 1
  sex_vector_average[row, ] <- (average_face - obs.vect[i,] ) *1.5
  row <- row + 1
}

#Get landmarks
new_means <- matrix(0, nrow = nrow(sex_vector_average), ncol = dim(eigenvec)[1])
  
for(i in 1:nrow(new_means)){
  new_means[i,] <- t(as.matrix(means))
}
  
landmarks <- (as.matrix(sex_vector_average) %*% t(as.matrix(eigenvec)) ) + new_means
#Saving dataset
setwd(paste(path, "/Results/Landmarks", sep = ""))
write_csv(as.data.frame(landmarks), "sex_vector_average.csv", col_names = FALSE)
```


